---
title: "Data Validation Documentation "
author: "Institution Effectiveness"
date: "7/2/2022"
output: html_document
theme: united
new_session: yes
---


# Purpose

The document shows examples and use cases of using the utValidateR package.
See github repository: https://github.com/dsu-effectiveness/utValidateR

### Loading Libraries

```{r echo = FALSE, fig.cap = 'A figure caption.'}
library(utHelpR)
library(utValidateR)
library(tidyverse)
library(gt)

# This function makes the data tables
# This function makes the data tables
mkd_gt_table <- function(df, title, subtitle) {
  gt(df) %>%
  tab_header(
    title = md(title),
    subtitle = md(subtitle)
    ) %>%
  tab_options(
    heading.align = "left",
    table.border.top.color = "white",
    column_labels.border.bottom.color = "black",
    table.align = "left",
    table.width = pct(100)
  ) %>%
    cols_align(
      align = "auto",
      columns = everything()
    ) %>%
    opt_css(
      css = "
      .gt_title {
                  font-size: 1.2em !important;
                  font-weight: 600 !important;
                }
      "
    )
}
```

## Buildings

### Buildings
- This code pulls the data
```{r echo = TRUE, warning=FALSE}

fake_buildings_df <- (utValidateR::fake_buildings_validation)
real_buildings_df <-(utHelpR::get_data_from_sql_file("buildings_validation.sql", "edify", "sandbox"))

mkd_gt_table(fake_buildings_df, "Fake Data Results:", "")
mkd_gt_table(real_buildings_df, "Real Data Results:", "")
```

- Checklist contains the validation rules and aux_info shows the valid value comparisons
```{r echo = TRUE, warning=FALSE}
data("checklist")
data("aux_info")
```

- Set a variable called building_checks which grabs the database validation rules for buildings
```{r echo = TRUE, warning=FALSE}
building_checks <-filter(checklist, file == "Buildings", type == "Database")
DT::datatable(
  building_checks,
  options = list(scrollX = TRUE),
  caption = "building_checks Results:"
)
```

- Run the do_checks function and be sure to pass in the data frame, checks, and aux_info

```{r echo = TRUE, warning=FALSE}
#Run do_checks and product validation results
building_validation_results <- do_checks(real_buildings_df, building_checks, aux_info)
  
DT::datatable(
  building_validation_results,
  options = list(scrollX = TRUE),
  caption = "building_validation_results Results:"
)

results <- building_validation_results %>% 
  select(ends_with("status")) %>% 
  setNames(gsub("_status", "", names(.))) %>% 
  summarize(across(.fns = function(x) mean(x == "Pass"))) %>% 
  pivot_longer(cols = everything(), names_to = "rule", values_to = "pass_rate") %>% 
  arrange(pass_rate) %>% 
  left_join(checklist, by = "rule")
```
