---
title: "S_14"
output: html_document
date: '2022-06-17'
---

```{r set global chunks, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE, message = FALSE
)
```

### Markdown libraries

Here we list all the libraries needed for this markdown document. The functions listed here are not dependent upon these libraries.

```{r}
library(DT)
library(purrr)
library(here)
library(stringr)
```


### Function dependent libraries

These libraries should go in the package dependency list. 

```{r function dependent libraries}
library(dplyr)
library(rlang)
```


### Fake data sets

```{r}
fake_df <- readRDS(here::here('sandbox', 'student', 'fake.RDS'))

DT::datatable(head(fake_df))
```

```{r}
s_14_test_values <- head(fake_df, 20) %>% 
  mutate(birth_date = ifelse(as.integer(student_id) %% 2 == 0, NA, birth_date)) 

DT::datatable(s_14_test_values)
```

## S_14_missing_birthdate

We want this function to append three columns to the input data frame. 

- `s_14_status`: This is a boolean indicating is the birth date is present or not.
- `s_14_activity_date`: This is a date field pulled from spbpers indicating the minimum date birth date was last updated.
- `s_14_age`: Integer field giving the minimum age of the birth date. 

```{r}
s_14_missing_birthdate <- s_14_test_values %>% 
  mutate(s_14_status = is.na(birth_date)) %>% 
  mutate(s_14_activity_date = spbpers_activity_date) %>% 
  mutate(s_14_age = spbpers_activity_date_days_past) 
```

```{r}
datatable(s_14_missing_birthdate)
```

### The function

I am imagining we are going to have several functions which test for missing elements, so let's just put together a single function to help with all of them.

```{r}
check_missing_field <- function(input_df, input_field_name, activity_date_name, output_field_name) {
  
  # Capture the output field name as a text element so that we can use it in
  # in the mutate statement as the name of the field.
  ofn_char <- ensym(output_field_name)
  
  ofn <- paste0(ofn_char, '_status')
  activity_char <- paste0(ofn_char, '_activity_date')
  
  output_df <- input_df %>% 
    mutate( {{ofn}} := is.na( {{input_field_name}} )) %>% 
    mutate( {{activity_char}} := {{activity_date_name}}   ) 
  
  return(output_df)
  
}
```

```{r}
check_missing_field(s_14_test_values, birth_date, spbpers_activity_date, s_14) %>% View()
```



```{r}
check_missing_field(input_df, birth_date, spbpers_activity_date, s_14)
```



