---
title: "S_14"
output: html_document
date: '2022-06-17'
---

```{r set global chunks, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE, message = FALSE
)
```

### Markdown libraries

Here we list all the libraries needed for this markdown document. The functions listed here are not dependent upon these libraries.

```{r}
library(DT)
library(purrr)
library(here)
library(stringr)
library(utValidateR)
```


### Function dependent libraries

These libraries should go in the package dependency list. 

```{r function dependent libraries}
library(dplyr)
library(rlang)
library(lubridate)
```


### Fake data sets

```{r}
data("fake_student_df")
s_14_test_values <- fake_student_df[21:30,] %>% 
  mutate(birth_date = NA) %>% 
  bind_rows(head(fake_student_df, 20))

DT::datatable(s_14_test_values)
```

## S_14_missing_birthdate

### Getting the age of an error

We will use the `lubridate` package to do this.

```{r}
test_date <- s_14_test_values$spriden_activity_date[1]
interval(test_date, Sys.Date()) %/% days(x=1)
```

The test date is `r test_date` and the system date is `r Sys.Date()`. 

```{r}
get_age <- function(x) {
  interval(x, Sys.Date()) %/% days(x=1)
}
```


### Building the function

We want this function to append three columns to the input data frame. 

- `s_14_status`: This is a boolean indicating if the birth date is present or not.
- `s_14_activity_date`: This is a date field pulled from spbpers indicating the minimum date birth date was last updated.
- `s_14_age`: Integer field giving the minimum age of the birth date. 

```{r}
s_14_missing_birthdate <- s_14_test_values %>% 
  mutate(s_14_status = is.na(birth_date)) %>% 
  mutate(s_14_activity_date = spriden_activity_date) %>% 
  mutate(s_14_error_age = get_age(spriden_activity_date))
```

```{r}
DT::datatable(s_14_missing_birthdate)
```

### The functions

I am imagining we are going to have several functions which test for missing elements, so let's just put together a single function to help with all of them. We will also need to add `get_age()` to the package.

```{r}
check_missing_field <- function(input_df, input_field_name, activity_date_name, output_field_name) {
  
  # Capture the output field name as a text element so that we can use it in
  # in the mutate statement as the name of the field.
  ofn_char <- ensym(output_field_name)
  
  ofn <- paste0(ofn_char, '_status')
  activity_char <- paste0(ofn_char, '_activity_date')
  age_char <- paste0(ofn_char, '_error_age')
  
  output_df <- input_df %>% 
    mutate( {{ofn}} := is.na( {{input_field_name}} )) %>% 
    mutate( {{activity_char}} := {{activity_date_name}}   ) %>% 
    mutate( {{age_char}} := get_age( {{activity_date_name}}  )     )
  
  return(output_df)
  
}
```

```{r}
check_missing_field(s_14_test_values, birth_date, spriden_activity_date, s_14) %>% 
  DT::datatable()
```




